---
title: "Weekly chart data analysis"
author: "Gregor Aisch"
date: "9/22/2020"
output:
  html_document: 
    df_print: tibble
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Load some useful packages using `needs`

```{r}
needs(tidyverse, directlabels, ggrepel)
```

Let's load yearly, seasonal and monthly averages for air temperature

```{r temp, echo=F, message=FALSE, warning=FALSE, cache=TRUE}
temp <- read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/annual/air_temperature_mean/regional_averages_tm_year.txt',
                   delim=';', skip = 1) %>% 
  transmute(year=Jahr, time='Year', value=as.numeric(Deutschland)) %>% 
  bind_rows(
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/seasonal/air_temperature_mean/regional_averages_tm_autumn.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='Autumn', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/seasonal/air_temperature_mean/regional_averages_tm_spring.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='Spring', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/seasonal/air_temperature_mean/regional_averages_tm_summer.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='Summer', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/seasonal/air_temperature_mean/regional_averages_tm_winter.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='Winter', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_01.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='January', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_02.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='February', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_03.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='March', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_04.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='April', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_05.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='May', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_06.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='June', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_07.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='July', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_08.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='August', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_09.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='September', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_10.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='October', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_11.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='November', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/air_temperature_mean/regional_averages_tm_12.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='December', value=as.numeric(Deutschland))
  ) %>% 
  mutate(measure='temperature')

temp
```

Load yearly, seasonal and monthly averages for precipitation
 
```{r precip, echo=F, message=FALSE, warning=FALSE, cache=TRUE}
precip <- read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/annual/precipitation/regional_averages_rr_year.txt',
                     delim=';', skip = 1) %>% 
  transmute(year=Jahr, time='Year', value=as.numeric(Deutschland)) %>% 
  bind_rows(
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/seasonal/precipitation/regional_averages_rr_autumn.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='Autumn', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/seasonal/precipitation/regional_averages_rr_spring.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='Spring', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/seasonal/precipitation/regional_averages_rr_summer.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='Summer', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/seasonal/precipitation/regional_averages_rr_winter.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='Winter', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_01.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='January', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_02.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='February', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_03.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='March', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_04.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='April', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_05.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='May', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_06.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='June', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_07.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='July', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_08.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='August', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_09.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='September', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_10.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='October', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_11.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='November', value=as.numeric(Deutschland)),
    read_delim('https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/monthly/precipitation/regional_averages_rr_12.txt',
               delim=';', skip = 1) %>% 
      transmute(year=Jahr,time='December', value=as.numeric(Deutschland))
  ) %>% 
  mutate(measure='precipitation')

precip
```

compute climate baseline (1961-1990)

```{r}
temp.base <- temp %>% 
  filter(year>1960 & year <= 1990) %>% 
  group_by(time) %>% 
  summarise(base=mean(value))
temp.base
```

```{r}
precip.base <- precip %>% 
  filter(year>1960 & year <= 1990) %>% 
  group_by(time) %>% 
  summarise(base=mean(value))
precip.base
```

Join temperature and precipitation data, add baseline, compute anomalies.

```{r out, message=FALSE, cache=TRUE}
out <- temp %>% 
  left_join(temp.base) %>% 
  bind_rows(left_join(precip, precip.base)) %>%            
  mutate(anomaly=ifelse(measure=='temperature',
                        value-base,             # difference anomaly for temp.
                        (value-base)/base*100)) %>%  # pct. anomaly for rain 
  select(year, time, measure, anomaly) %>% 
  pivot_wider(names_from=measure, values_from=anomaly)

out
```

Save the 2019 & 2020 data for our first plot.

```{r}
out %>% 
  filter(year==2019 & !(time %in% c('Year','Spring','Summer','Winter','Autumn'))) %>% 
  write_csv('temp-precip-anomalies-2019.csv')

out %>% 
  filter(year==2020 & !(time %in% c('Year','Spring','Summer','Winter','Autumn'))) %>% 
  write_csv('temp-precip-anomalies-2020.csv')
```

For the scatterplot custom line annotations we need to generate some markup:

```{r}
colors <- tribble(#####1976b3
  ~time, ~color,
  'Year', '#333333',
  'Summer', '#308c00',
  'June', '#308c00',
  'July', '#308c00',
  'August', '#308c00',
  #~~~~~~~~~~~~~~~~~
  'Winter', '#6ea2ff',
  'December', '#6ea2ff',
  'January', '#6ea2ff',
  'February', '#6ea2ff',
  #~~~~~~~~~~~~~~~~~
  'Autumn', '#ac2125',
  'September', '#ac2125',
  'October', '#ac2125',
  'November', '#ac2125',
  #~~~~~~~~~~~~~~~~~
  'Spring', '#fac10e',
  'March', '#fac10e',
  'April', '#fac10e',
  'May', '#fac10e',
)

out %>% 
  filter(year==2020) %>% 
  filter(!(time %in% c('Year','Summer','Winter','Spring','Autumn'))) %>% 
  left_join(colors) %>% 
  transmute(x1=0,
            y1=0,
            x2=round(temperature,2),
            y2=round(precipitation,2),
            paste0('@color:',color),
            '@width:2') %>% 
  format_csv(col_names = F) %>% 
  str_replace_all(',@', ' @')
```

### Time for some plotting

Plot all months for 2019

```{r plot-2019, fig.width=7, fig.height=7, cache=T}
out %>% 
  filter(year==2019) %>% 
  filter(!(time %in% c('Summer','Winter','Spring','Autumn', 'Year'))) %>% 
  ggplot(aes(xend=temperature, yend=precipitation,color=time)) +
  geom_hline(yintercept=0)  +
  geom_vline(xintercept=0)  +
  geom_segment(x=0,y=0, arrow = arrow(length = unit(7, 'points'))) +
  geom_text(aes(temperature, precipitation, label=time)) +
  theme_minimal()
```
Plot all months for 2020

```{r plot-2020, fig.width=7, fig.height=7, cache=T}
out %>% 
  filter(year==2020) %>% 
  filter(!(time %in% c('Summer','Winter','Spring','Autumn', 'Year'))) %>% 
  ggplot(aes(xend=temperature, yend=precipitation,color=time)) +
  geom_hline(yintercept=0)  +
  geom_vline(xintercept=0)  +
  geom_segment(x=0,y=0, arrow = arrow(length = unit(7, 'points'))) +
  geom_text(aes(temperature, precipitation, label=time)) +
  theme_minimal()
```
Plot all seasons for years since 2010

```{r seasons, fig.width=7, fig.height=7, cache=T}
d <- out %>% 
  filter(year>=2010) %>% 
  filter(time %in% c('Summer','Winter','Spring','Autumn'))
  
d %>% write_csv('temp-precip-anomalies-2010-2020.csv')

d %>% 
  ggplot(aes(xend=temperature, yend=precipitation,color=time)) +
  geom_hline(yintercept=0)  +
  geom_vline(xintercept=0)  +
  geom_segment(x=0,y=0, arrow = arrow(length = unit(7, 'points'))) +
  geom_text(aes(temperature, precipitation, label=year)) +
  theme_minimal()
```

Average seasons since 2000:

```{r avg-seasons, fig.width=7, fig.height=7, cache=T}
out %>% 
  filter(year>=2000) %>% 
  filter(time %in% c('Summer','Winter','Spring','Autumn')) %>% 
  group_by(time) %>% 
  summarise(temperature=mean(temperature), precipitation=mean(precipitation)) %>% 
  ggplot(aes(xend=temperature, yend=precipitation,color=time)) +
  geom_hline(yintercept=0)  +
  geom_vline(xintercept=0)  +
  geom_text_repel(aes(temperature, precipitation, label=time)) +
  theme_minimal() +
  geom_segment(x=0,y=0, arrow = arrow(length = unit(7, 'points')))

```

Let's look at where all the years since 1961 are in the plot (highlighting the most recent years):

```{r all_years, echo=FALSE, fig.width=7, fig.height=7}
out %>% 
  filter(time=='Year' & year > 1960) %>% 
  mutate(recent=year>=2010) %>% 
  ggplot(aes(temperature, precipitation, label=year, color=recent)) +
  geom_point() +
  geom_text_repel() +
  geom_hline(yintercept=0)  +
  geom_vline(xintercept=0)  +
  theme_minimal()
```


